Shrinking Characteristics of Precision Matrix Estimators: An Illustration via Regression

### 2019-04-20 05-appendix.Rmd
complete appendix near final

@witten2009covariance and @price2015ridge showed that the following optimization problem can be solved in closed-form. We outline their steps below.

@@

\begin{equation}
\begin{split}
  \hat{Omega}^{k + 1} &:= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\tilde{L}_{\rho}(\Omega, Z^{k}, \Lambda^{k}) \\
  &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{\begin{matrix}
 tr\left(S\Omega\right) - \log\left|\Omega\right| + tr\left[(\Lambda^{k})^{T}(A\Omega B - Z^{k} - C) \right] + \rho\left\|A\Omega B - Z^{k} - C \right\|_{F}^{2}/2 \end{matrix}\right. \\
  &+ \left.\begin{matrix} vec\left(\Omega - \Omega^{k}\right)^{T}\rho Q\left(\Omega - \Omega^{k}\right)/2 \end{matrix}\right\} \\
  &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{\begin{matrix}
 tr\left(S\Omega\right) - \log\left|\Omega\right| + tr\left[(\Lambda^{k})^{T}(A\Omega B - Z^{k} - C) \right] + \rho\left\|A\Omega^{k} B - Z^{k} - C \right\|_{F}^{2}/2 \end{matrix}\right. \\
  &+ \left.\begin{matrix} \rho\tau\left\|\Omega - \Omega^{k}\right\|_{F}^{2}/2 + tr\left[\rho\left(\Omega - \Omega^{k}\right)\left(BB^{T}\Omega^{k}A^{T}A - B(Z^{k})^{T}A - BC^{T}A \right)\right] \end{matrix}\right\} \\
  &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{\begin{matrix}
 tr\left[\left(S + \rho A^{T}(A\Omega^{k}B - Z^{k} - C + \Lambda^{k}/\rho)B^{T} \right)\Omega\right] \end{matrix}\right. \\
  &- \left.\begin{matrix} \log\left|\Omega\right| + \rho\tau\left\|\Omega - \Omega^{k}\right\|_{F}^{2}/2 \end{matrix}\right\} \\
  &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{
 tr\left[\left(S + G^{k} \right)\Omega\right] - \log\left|\Omega\right| + \rho\tau\left\|\Omega - \Omega^{k}\right\|_{F}^{2}/2 \right\} \\
(\#eq:omegaapprox)\notag
\end{split}
\end{equation}

@@

\begin{equation}
Z_{ij}^{k + 1} = \frac{1}{\rho}\left( \rho(A\Omega_{ij}^{k + 1}B - C) + \Lambda_{ij}^{k} - Sign(Z_{ij}^{k + 1})\lambda \right)
(\#eq:ZZsoftproof2)\notag
\end{equation}

@@

This implies that $\mbox{sign}(Z_{ij}^{k + 1}) = \mbox{sign}\left(\rho(A\Omega_{ij}^{k + 1}B - C) + \Lambda_{ij}^{k}\right)$. Putting all the pieces together, we arrive at

\begin{equation}
\begin{split}
Z_{ij}^{k + 1} &= \frac{1}{\rho}\mbox{sign}\left(\rho(A\Omega_{ij}^{k + 1}B - C) + \Lambda_{ij}^{k}\right)\left( \left| \rho(A\Omega_{ij}^{k + 1}B - C) + \Lambda_{ij}^{k} \right| - \lambda \right)_{+} \\
&= \frac{1}{\rho}\mbox{soft}\left(\rho(A\Omega_{ij}^{k + 1}B - C) + \Lambda_{ij}^{k}, \lambda\right)
(\#eq:appendixZZproof2)\notag
\end{split}
\end{equation}

>>>>

@witten2009covariance and @price2015ridge showed that the following optimization problem can be solved in closed-form. We outline their steps here.

@@

\begin{equation}
\begin{split}
  \hat{\Omega}^{k + 1} &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\tilde{L}_{\rho}(\Omega, Z^{k}, \Lambda^{k}) \\
  &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{\begin{matrix}
 tr\left(S\Omega\right) - \log\left|\Omega\right| + tr\left[(\Lambda^{k})^{T}(A\Omega B - Z^{k} - C) \right] + \rho\left\|A\Omega B - Z^{k} - C \right\|_{F}^{2}/2 \end{matrix}\right. \\
  &+ \left.\begin{matrix} vec\left(\Omega - \Omega^{k}\right)^{T}\rho Q\left(\Omega - \Omega^{k}\right)/2 \end{matrix}\right\} \\
  &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{\begin{matrix}
 tr\left(S\Omega\right) - \log\left|\Omega\right| + tr\left[(\Lambda^{k})^{T}(A\Omega B - Z^{k} - C) \right] + \rho\left\|A\Omega^{k} B - Z^{k} - C \right\|_{F}^{2}/2 \end{matrix}\right. \\
  &+ \left.\begin{matrix} \rho\tau\left\|\Omega - \Omega^{k}\right\|_{F}^{2}/2 + tr\left[\rho\left(\Omega - \Omega^{k}\right)\left(BB^{T}\Omega^{k}A^{T}A - B(Z^{k})^{T}A - BC^{T}A \right)\right] \end{matrix}\right\} \\
  &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{\begin{matrix}
 tr\left[\left(S + \rho A^{T}(A\Omega^{k}B - Z^{k} - C + \Lambda^{k}/\rho)B^{T} \right)\Omega\right] \end{matrix}\right. \\
  &- \left.\begin{matrix} \log\left|\Omega\right| + \rho\tau\left\|\Omega - \Omega^{k}\right\|_{F}^{2}/2 \end{matrix}\right\} \\
  &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{
 tr\left[\left(S + G^{k} \right)\Omega\right] - \log\left|\Omega\right| + \rho\tau\left\|\Omega - \Omega^{k}\right\|_{F}^{2}/2 \right\} \\
(\#eq:omegaapprox)\notag
\end{split}
\end{equation}

@@

\begin{equation}
Z_{ij}^{k + 1} = \frac{1}{\rho}\left( \rho\left(A\Omega_{ij}^{k + 1}B - C\right) + \Lambda_{ij}^{k} - Sign\left(Z_{ij}^{k + 1}\right)\lambda \right)
(\#eq:ZZsoftproof2)\notag
\end{equation}

@@

This implies that $\mbox{sign}\left(Z_{ij}^{k + 1}\right) = \mbox{sign}\left(\rho\left(A\Omega_{ij}^{k + 1}B - C\right) + \Lambda_{ij}^{k}\right)$. Putting all the pieces together, we arrive at

\begin{equation}
\begin{split}
Z_{ij}^{k + 1} &= \frac{1}{\rho}\mbox{sign}\left(\rho\left(A\Omega_{ij}^{k + 1}B - C\right) + \Lambda_{ij}^{k}\right)\left( \left| \rho\left(A\Omega_{ij}^{k + 1}B - C\right) + \Lambda_{ij}^{k} \right| - \lambda \right)_{+} \\
&= \frac{1}{\rho}\mbox{soft}\left(\rho\left(A\Omega_{ij}^{k + 1}B - C\right) + \Lambda_{ij}^{k}, \lambda\right)
(\#eq:appendixZZproof2)\notag
\end{split}
\end{equation}
